A Logistic Regression Approach to CoIL Challenge 2000

### 2019-12-15 final project/Draft.Rmd

---
title             : "A Logistic Regression Approach to CoIL Challenge 2000"
shorttitle        : " "
author            : "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
affiliation       : "MS in Data Science Students CUNY School of Professional Studies"
abstract: |
  A logistic regression based solution to the CoIL Challenge 2000 is described. The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
  <!-- https://tinyurl.com/ybremelq 
    One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  -->
keywords          : "CoIL Challenge, Logistic Regression"
wordcount         : "X"
bibliography      : ["Manuscript.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output: 
  word_document:
    reference_docx: word-template.docx
  papaja::apa6_pdf:
    #citation_package: natbib #biblatex
  html_document: default
---

>>>>

---
title             : "A Logistic Regression Approach to CoIL Challenge 2000"
shorttitle        : " "
author            : "Corey Arnouts, Adam Douglas, Jason Givens-Doyle, Michael Silva"
affiliation       : "MS in Data Science Students CUNY School of Professional Studies"
abstract: |
  A logistic regression based solution to the CoIL Challenge 2000 is described. The challenge consists of correctly identifying potential customers for an insurance product, and describing their characteristics.
keywords          : "CoIL Challenge, Logistic Regression"
wordcount         : "X"
bibliography      : ["Manuscript.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output: 
  word_document:
    reference_docx: word-template.docx
  papaja::apa6_pdf:
    #citation_package: natbib #biblatex
  html_document: default
---

### 2019-12-15 final project/Manuscript.Rmd

```{r echo=FALSE, message=FALSE}
temp <- read.delim("data/ticdata2000.txt", header = FALSE, nrows = 1)
df_cols <- length(names(temp))
source("CoIL.R")
library(pander)
knitr::opts_chunk$set(comment = NA)
knitr::knit_hooks$set(inline = function(x) { if(!is.numeric(x)){ x }else{ prettyNum(round(x,2), big.mark=",") } })
```

@@

```{r, echo=FALSE}
# Similar the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121), n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
ggplot(data = prediction_scores, aes(x = score, y = n)) +
  geom_bar(stat = "identity", fill = "#377eb8") +
  xlab("Number of Correct Predictions") +
  ylab(element_blank()) + 
  xlim(37, 240) +
  annotate("text", x = 65, y = 4, label = "Previous Researchers") +
  geom_vline(xintercept = 238, color = "#e41a1c", linetype="dashed") +
  annotate("text", x = 220, y = 4, label = "Maximum") +
  geom_vline(xintercept = final_model$correct, color = "#4daf4a") +
  annotate("text", x = final_model$correct - 20, y = 4, label = "Our Model") +
  ggtitle("Number of Customers Correctly Predicted")
```

The maximum number of potential policy owners that could be found is 238.  Previous researchers identified 95 policy owners on average.  The best preforming model [@Elkan] during the initial challenge identified 121 policy owners.  A meta analysis found that simpler algorithms tended to outpreform more complicated ones [@Putten].

@@

<!-- Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection. -->
The CoIL Challenge dataset is composed of `r df_cols` variables accross `r nrow(df)` observations.  An evaluation dataset is provided with `r nrow(eval)` observations.

**THERE ARE X that are categorical and X that are numeric.**  Most of the variables have little to no correlation with the variable of interest (CARAVAN).

@@

The MOSTYPE variable has 40 customer types.  Not all customer types were statiscally significant predictors.  We identified those with a probability greater than 0.5 to purchase the insurance product as the `LIKELY_CUSTOMER` variable.  Our second model used LIKELY_CUSTOMER and PPERSAUT.

@@

```{r, echo=FALSE, fig.width=10, fig.height=11}
temp <- up_train %>%
  group_by(MOSTYPE) %>%
  tally() %>%
  rename(total = n)

@@

## Correlation Coefficients for Variables of Interest
```{r, echo=FALSE}
vars <- c("CARAVAN", "PPERSAUT", "PBRAND", "APERSAUT", "LIKELY_CUSTOMERS")
temp <- up_train %>%
  mutate(CARAVAN = as.numeric(as.character(CARAVAN)),
         LIKELY_CUSTOMERS = as.numeric(as.character(LIKELY_CUSTOMERS)))
temp <- temp[vars]

pander(cor(temp[vars], temp[vars]))
```

## Data Dictionary

```{r, results='asis', echo=FALSE}

@@

```{r, echo=FALSE}
final_model$cm
```

@@

```{r, null_prefix = TRUE, comment=NA, echo=FALSE}
conn <- file("CoIL.r", open = "r")
lines <- readLines(conn)
n_lines <- length(lines)
for (i in 1:n_lines) {
  cat(lines[i], "\n")
}
close(conn)
```

>>>>

```{r echo=FALSE, message=FALSE}
temp <- read.delim("data/ticdata2000.txt", header = FALSE, nrows = 1)
df_cols <- length(names(temp))
source("CoIL.R")
library(pander)
knitr::opts_chunk$set(comment = NA, echo = FALSE)
knitr::knit_hooks$set(inline = function(x) { if(!is.numeric(x)){ x }else{ prettyNum(round(x,2), big.mark=",") } })
```

@@

The maximum number of potential policy owners that could be found is 238.  Previous researchers identified 95 policy owners on average.  The best preforming model [@Elkan] during the initial challenge identified 121 policy owners.  It was a Naïve Bayes, suggesting that probabilities of some of the variables will be useful in identifing potential customers.

A meta analysis of the initial researchers found that simpler algorithms tended to outpreform more complicated ones [@Putten].  With the benefit of these findings, we set out to create a simple logistic regression model that preforms as well or better than the original CoIL Challenge cohort.  In the end, our model outpreformed the orignal researcher's model in correctly predicting the customers that would purchase the insurance policy.

```{r}
# Similar the chart found on page 2 of http://liacs.leidenuniv.nl/~puttenpwhvander/library/cc2000/PUTTEN~1.pdf
prediction_scores <- data.frame(score = c(38, 46, 53, 58, 65, 72, 74, 80, 83, 86, 91, 94, 95, 96, 97, 99, 101, 103, 104, 105, 106, 107, 108, 109, 110, 111, 112, 115, 121), n = c(1, 1, 1, 1, 1, 1, 1, 1, 1, 2, 1, 1, 1, 3, 1, 2, 1, 1, 1, 3, 1, 2, 2, 4, 2, 2, 2, 1, 1))
ggplot(data = prediction_scores, aes(x = score, y = n)) +
  geom_bar(stat = "identity", fill = "#377eb8") +
  xlab(element_blank()) +
  ylab(element_blank()) + 
  xlim(37, 250) +
  annotate("text", x = 65, y = 4, label = "Previous Researchers") +
  geom_vline(xintercept = 238, color = "#e41a1c", linetype="dashed") +
  annotate("text", x = 220, y = 4, label = "Maximum") +
  geom_vline(xintercept = final_model$correct, color = "#4daf4a") +
  annotate("text", x = final_model$correct - 20, y = 4, label = "Our Model") +
  ggtitle("Number of Customers Correctly Predicted")
```

@@

<!-- Discuss the key aspects of your problem, data set and regression model(s). Given that you are working on real-world data, explain at a high-level your exploratory data analysis, how you prepared the data for regression modeling, your process for building regression models, and your model selection. -->
The CoIL Challenge dataset is composed of `r df_cols` variables accross `r nrow(df)` observations.  An evaluation dataset is provided with `r nrow(eval)` observations. Five of the predictors are categorical and the remainder are numeric.  Most of the predictors have little to no correlation with the variable of interest (CARAVAN).

@@

The MOSTYPE variable has 40 customer types.  Not all customer types were statiscally significant predictors.  We identified those in the oversampled dataset with a probability greater than 0.5 to purchase the insurance product as the LIKELY_CUSTOMER variable.  A figure in the appendix shows all 40 customer types and the probability that they purched the insurance policy in the oversampled dataset.

We fit our second model to the LIKELY_CUSTOMER and PPERSAUT variables. Although it only has two variables, the model preformed well.  This seemed in-line with expectations that simple models preform best.

In evaluating the models we examined we focused on the specificity.  The goal of the CoIL challenge was to accurately predict those would would purchase the insurance policy, so focusing on the model's specificity was the best evaluation metric.  In order to get a better sense of how well the model generalizes, we repeatedly retrained and evaluated the model using diffrent samplings of the training dataset.  The following figure summarizes the distribution of the specificty the models produced on the test dataset:

```{r}
#temp <- rbind(data.frame("Model" = c("1"), "Specificity" = model1_robust_results$specificity),
#              data.frame("Model" = c("2"), "Specificity" = model2_robust_results$specificity))%>%
temp <- rbind(data.frame("Model" = c("2"), "Specificity" = model2_robust_results$specificity))%>%
  mutate(Model = as.factor(Model))

ggplot(temp, aes(x = Specificity, fill = Model, color = Model)) +
  geom_density(alpha = 0.5) +
  ylab("Density") +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  #geom_vline(xintercept = mean(model1_robust_results$specificity), linetype="dashed", color = "#e41a1c") +
  geom_vline(xintercept = mean(model2_robust_results$specificity), linetype="dashed", color = "#377eb8")
```

@@

```{r, fig.width=10, fig.height=11}
temp <- up_train %>%
  group_by(MOSTYPE) %>%
  tally() %>%
  rename(total = n)

@@

## Correlation Coefficients for Variables of Interest
```{r}
vars <- c("CARAVAN", "MOSTYPE", "PPERSAUT", "PBRAND", "MOSHOOFD", "APERSAUT", "LIKELY_CUSTOMERS")
temp <- up_train %>%
  mutate(CARAVAN = as.numeric(as.character(CARAVAN)),
         LIKELY_CUSTOMERS = as.numeric(as.character(LIKELY_CUSTOMERS)))
temp <- temp[vars]

numeric_vars <- unlist(lapply(temp, is.numeric))

pander(cor(temp[numeric_vars], temp[numeric_vars]))
```

## Data Dictionary for Variables of Interest

```{r, results='asis'}

@@

```{r}
final_model$cm
```

@@

```{r}
conn <- file("CoIL.r", open = "r")
lines <- readLines(conn)
n_lines <- length(lines)
for (i in 1:n_lines) {
  cat(lines[i], "\n")
}
close(conn)
```
